XOR-Some Prolific Pilot 2

### 2021-04-09 analysis/xor-some-prolific-pilot2.Rmd
finish code

Compute the average deviation from expected responses for each vignette, individually for each dimension (rel / comp / pri):
```{r}
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  mutate(expected_rating = prior_class * 100,
         response_deviation = abs(response - expected_rating)) %>%
  group_by(title, block_extended) %>%
  summarize(mean_deviation = mean(response_deviation)) %>%
  arrange(desc(mean_deviation), .by_group = T) %>%
  # get the largest deviations - over 50 
  filter(mean_deviation >= 50) -> d_critical_deviations

@@

# count which dimension is the most problematic - the prior
d_critical_deviations %>% group_by(block_extended) %>% count()

@@

@@

d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) # get ratings from the respective trials only

```
```{r}
# check whether the ratings in the two prior questions are the same in xor 
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  filter(block == "prior", main_type == "xor") %>%
  mutate(priorQ_nr = rep(c(1,2), 396)) -> d_xor_priors
d_xor_priors %>% group_by(priorQ_nr, prior_class) %>%
  summarise(mean = mean(response)) -> d_xor_priors_summary
d_xor_priors %>%
  ggplot(., aes(x = as.factor(priorQ_nr), y = response )) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
  xlab("First vs Second prior question for high vs low prior conditions") +
  facet_wrap(~as.factor(prior_class)) # get ratings from the respective trials only

@@

Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical_long %>% 
  filter(block == "xor" | block == "some") %>%
  #group_by(block, class_condition, prior_class) %>%
  #summarize(mean_rating = mean(response)) %>%
  ggplot(., aes(x = prior_class, y = response)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method="lm") +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Inference strength ratings by-predictor")

@@

Plot inference strength ratings against raw predictor ratings for each participant / each item:
```{r}
d_critical_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

@@

Plot mean ratings (across all participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition. For relevance and competence, the color indicates whether it was presented without the utterance or with. 
```{r, fig.width=10, fig.height=30}
bar.width = 0.8
d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_")) %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  group_by(block_extended, title, w_utterance, main_type) %>%
  summarize(mean_rating = mean(response)) %>% 
  ggplot(., aes(x = block_extended, y = mean_rating, fill = w_utterance)) +
  geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
  #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question mean ratings")
```

@@

# correlation plot for "some"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "some") %>%  
    select(prior:ncol(d_critical_wide))
)

# correlation plot for "or"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "xor") %>%  
    select(prior:ncol(d_critical_wide))
  )
```

Check by-participant behavior to see if they show versatile response strategies for the different questions for xor vs. some. Participants show a sufficient range of responses across conditions.  
```{r, fig.width=15, fig.height=7}
d_critical_long %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, size = 2) + 
  facet_wrap(main_type~submission_id, ncol = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-participant ratings all xor vs. some trials")
```

## Explore z-scoring
Visual exploration of z-scoring each measure (rel / comp / pri / target / comp-wUtterance / rel-wUtterance) per participant, collapsing some and or together. 
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd)

@@

d_critical_zScored %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response_centered, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_zScore_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) + # get ratings from the respective trials only 
  ggtitle("Centered responses to by-predictor questions")
```

@@

d_critical_zScored_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel.z

@@

@@

@@

@@

@@

#### Stats with dichotomized predictors 
Maximal model on raw inference ratings, using anticipated binary categorization of items as predictors 
```{r, results='hide'}
d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = as.factor(relevance),
         competence = as.factor(competence),
         prior = as.factor(prior)) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat
# sum code predictors
# low prior: 1, high prior:: -1 
contrasts(d_critical_wide_cat$prior) <- contr.sum(2)
# low comp : 1, high comp : -1
contrasts(d_critical_wide_cat$competence) <- contr.sum(2)
# low rel: 1, high rel: -1
contrasts(d_critical_wide_cat$relevance) <- contr.sum(2)

@@

```{r}
summary(model_xor_cat)
```

```{r, results='hide'}
model_some_cat <- brm(
  response ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some_cat)
```

Maximal model on z-scored inference ratings, using anticipated binary categorization of items as predictors 
```{r, results='hide'}
d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = as.factor(relevance),
         competence = as.factor(competence),
         prior = as.factor(prior)) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat_zScored
# sum code predictors
# low prior: 1, high prior:: -1 
contrasts(d_critical_wide_cat_zScored$prior) <- contr.sum(2)
# low comp : 1, high comp : -1
contrasts(d_critical_wide_cat_zScored$competence) <- contr.sum(2)
# low rel: 1, high rel: -1
contrasts(d_critical_wide_cat_zScored$relevance) <- contr.sum(2)

@@

```{r, results='hide'}
model_some_cat_zScored <- brm(
  response_centered ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some_cat_zScored)
```

In summary, the different models provide somewhat different results, specially when using raw ratings vs. binary predictors. Since interpreting the categorical models in relation to the plots might be not as straight-forward, I am not entirely sure whether the qualitative results match in all models. Using centered data seems to provide a little tighter estimates in the models; not sure where the slight qualitative differences come from.

>>>>

Compute the average deviation from expected responses for each vignette, individually for each dimension (rel / comp / pri):
```{r}
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  mutate(expected_rating = prior_class * 100,
         response_deviation = abs(response - expected_rating),
         block = ifelse(block == "xor" | block == "some", "target", block)
         ) %>%
  group_by(title, block) %>%
  summarize(mean_deviation = mean(response_deviation)) %>%
  arrange(desc(mean_deviation), .by_group = T) %>%
  # get the largest deviations - over 50 
  filter(mean_deviation >= 50) -> d_critical_deviations

@@

# count which dimension is the most problematic - the prior
d_critical_deviations %>% group_by(block) %>% count()

@@

Compute the standard deviation for each question for each item as an additional measure of spread in the participants' responses within-item:
```{r}
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  mutate(
         block = ifelse(block == "xor" | block == "some", "target", block)
         ) %>%
  group_by(title, block) %>%
  summarize(item_sd = sd(response)) %>%
  arrange(desc(item_sd), .by_group = T) %>%
  # arbitrary set to 25
  filter(item_sd >= 25) -> d_critical_sd
d_critical_sd

#d_critical_sd %>% write_csv("../data/pilots/pilot2_byItem_rating_SDs.csv")
```

Visually check the spread of by-item ratings among subjects for each question:
```{r, fig.height=20, fig.width=7}

d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  filter(block == class_condition | block == "target") %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
  ylab("Raw responses to respective questions") +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question raw ratings")
```
For the following items, participants seem  to *disagree* in their judgments for all of the predictors: Emails from a broken laptop, Gift unwrapping, Martha's cookies, Suzie's fruits, Brad's clothes, Crime, maybe more..

Check by-vignette by-predictor empirical means, to see if the range is somewhat covered:
```{r}
d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  filter(block == class_condition | block == "target") %>%
  group_by(title, block) %>%
  summarise(mean_response = mean(response)) %>%
  arrange(mean_response)
```

@@

d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  ylab("Responses to respective predictor question") +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) # get ratings from the respective trials only

```
```{r}
# check whether the ratings in the two prior questions are the same in xor 
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  filter(block == "prior", main_type == "xor") %>%
  mutate(priorQ_nr = rep(c(1,2), 396)) -> d_xor_priors
d_xor_priors %>% group_by(priorQ_nr, prior_class) %>%
  summarise(mean = mean(response)) -> d_xor_priors_summary
d_xor_priors %>%
  ggplot(., aes(x = as.factor(priorQ_nr), y = response )) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
  ylab("Responses to prior questions") +
  xlab("First vs Second prior question for high vs low prior conditions") +
  facet_wrap(~as.factor(prior_class)) # get ratings from the respective trials only

@@

Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical_long %>% 
  filter(block == "xor" | block == "some") %>%
  ggplot(., aes(x = prior_class, y = response)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method="lm") +
  ylab("Inference strength ratings") +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Inference strength ratings by-predictor")

@@

Plot inference strength ratings against raw predictor ratings for each participant / each item:
```{r}
d_critical_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("Inference strength ratings") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

@@

Plot mean ratings (across all participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition. For relevance and competence, the color indicates whether it was presented without the utterance or with. 
```{r, fig.width=10, fig.height=30}
bar.width = 0.8
d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_")) %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  group_by(block_extended, title, w_utterance, main_type) %>%
  summarize(mean_rating = mean(response)) %>% 
  ggplot(., aes(x = block_extended, y = mean_rating, fill = w_utterance)) +
  geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
  #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
  ylab("Mean responses to respective questions") +
  xlab("Question type") +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question mean ratings")
```

Plot by-item trends between predictor ratings and target inference strength ratings (relevance / competence ratings with utterance are omitted):
```{r, fig.height=20, fig.width=8}
# make a versatile wide representation of the critical data
d_critical_predictorsXtarget <- d_critical_wide %>% 
  pivot_longer(
    cols = c(relevance, prior, competence, relevance_wUtt, competence_wUtt),
    names_to = "predictors",
    values_to = "predictor_response"
  ) %>%
  filter(predictors != "relevance_wUtt", predictors != "competence_wUtt")

d_critical_predictorsXtarget %>%
  ggplot(., aes(x = predictor_response, y = target)) +
  geom_point() +
  facet_wrap(title~predictors, ncol= 6) +
  xlab("Response to predictor questions") +
  ylab("Inference strength rating") +
  ggtitle("By-item Inference Strength Ratings vs. Predictor Ratings (left to right)")

Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants. Error bard indicate the standard deviation:
```{r, fig.height=4, fig.width=8}
d_critical_predictorsXtarget_summary <- d_critical_predictorsXtarget %>% 
  group_by(title, predictors, main_type) %>%
  summarise(mean_target = mean(target),
            mean_predictor = mean(predictor_response), 
            sd_target = sd(target),
            sd_predictor = sd(predictor_response))

d_critical_predictorsXtarget_summary %>%
  ggplot(., aes(x = mean_predictor, y = mean_target)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_target - sd_target, ymax = mean_target + sd_target), alpha = 0.7) +
  geom_errorbarh(aes(xmin = mean_predictor - sd_predictor, xmax = mean_predictor + sd_predictor), alpha = 0.7) +
  geom_text(aes(label=title), size = 1.5, nudge_y = 4) +
  geom_smooth(method = "lm") +
  ylab("Mean inference strength rating") +
  xlab("Mean predictor ratings") + 
  facet_wrap(main_type~predictors)
```

@@

# correlation plot for "some"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "some") %>%  
    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt) 
)

# correlation plot for "or"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "xor") %>%  
    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt)
  )
```

## Explore z-scoring
Visual exploration of z-scoring each measure (rel / comp / pri / target / comp-wUtterance / rel-wUtterance) per participant, collapsing some and or together. 
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))

@@

d_critical_zScored %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response_centered, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_zScore_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  ylab("Responses to respective predictor questions") +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) + # get ratings from the respective trials only 
  ggtitle("Centered responses to by-predictor questions")
```

@@

d_critical_zScored_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  ylab("Inference strength ratings") +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel.z

@@

Correlation plots of z-scored data:
``` {r}

# correlation plot for "some"
GGally::ggpairs(
  filter(d_critical_zScored_wide, main_type == "some") %>% ungroup() %>%  
    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt)
)

# correlation plot for "or"
GGally::ggpairs(
  filter(d_critical_zScored_wide, main_type == "xor") %>%  ungroup() %>%
    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt)
  )

@@

Exploring plots of posterior samples for pairwise parameters shows that there is a visible correlation between prior & competence, and a somewhat weaker correlation between competence & relevance. The interaction estimates also show a correlation with the main effects (as a rough overall judgment).   
```{r}
shinystan::launch_shinystan(model_xor)
```
Plot correlations of main effects
```{r}
bayesplot::mcmc_pairs(model_xor, pars = c("b_prior", "b_competence", "b_relevance"))
```

@@

For "some", the correlation of prior and relevance is higher than for prior and competence; there is also a visible correlation of competence and relevance. 
```{r}
shinystan::launch_shinystan(model_some)
```
```{r}
bayesplot::mcmc_pairs(model_some, pars = c("b_prior", "b_competence", "b_relevance"))
```

#### Smaller models with raw data to check for colinearity 
Models on raw data with only one predictor: 
```{r, results='hide'}
# xor
model_xor_prior <- brm(
  target ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_prior)
```

```{r, results='hide'}
# xor
model_xor_comp <- brm(
  target ~ competence + 
    (1 + competence | submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_comp)
```

```{r, results='hide'}
# xor
model_xor_rel <- brm(
  target ~ relevance + 
    (1 + relevance | submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_rel)
```

```{r, results='hide'}
# some
model_some_rel <- brm(
  target ~ relevance + 
    (1 + relevance | submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_rel)
```

```{r, results='hide'}
# some
model_some_prior <- brm(
  target ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_prior)
```

```{r, results='hide'}
# some
model_some_comp <- brm(
  target ~ competence + 
    (1 + competence | submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_comp)
```

With two predictors:
```{r, results='hide'}
# xor
model_xor_pri_comp <- brm(
  target ~ prior*competence + 
    (1 + prior*competence || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_pri_comp)
```

```{r}
bayesplot::mcmc_pairs(model_xor_pri_comp, pars = c("b_prior", "b_competence"))
```

```{r, results='hide'}
# xor
model_xor_pri_rel <- brm(
  target ~ prior*relevance + 
    (1 + prior*relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_pri_rel)
```

```{r}
bayesplot::mcmc_pairs(model_xor_pri_rel, pars = c("b_prior", "b_relevance"))
```

```{r, results='hide'}
# xor
model_xor_rel_comp <- brm(
  target ~ relevance*competence + 
    (1 + relevance*competence || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_rel_comp)
```

```{r}
bayesplot::mcmc_pairs(model_xor_rel_comp, pars = c("b_competence", "b_relevance"))
```

```{r, results='hide'}
# some
model_some_pri_comp <- brm(
  target ~ prior*competence + 
    (1 + prior*competence || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_pri_comp)
```

```{r}
bayesplot::mcmc_pairs(model_some_pri_comp, pars = c("b_prior", "b_competence"))
```

```{r, results='hide'}
# some
model_some_pri_rel <- brm(
  target ~ prior*relevance + 
    (1 + prior*relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_pri_rel)
```

```{r}
bayesplot::mcmc_pairs(model_some_pri_rel, pars = c("b_prior", "b_relevance"))
```

```{r, results='hide'}
# some
model_some_rel_comp <- brm(
  target ~ relevance*competence + 
    (1 + relevance*competence || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_rel_comp)
```

```{r}
bayesplot::mcmc_pairs(model_some_rel_comp, pars = c("b_competence", "b_relevance"))
```

@@

```{r}
bayesplot::mcmc_pairs(model_xor_zScored, pars = c("b_prior", "b_relevance", "b_competence"))
```

@@

```{r}
bayesplot::mcmc_pairs(model_some_zScored, pars = c("b_prior", "b_relevance", "b_competence"))
```

#### Smaller models with z-scored data
```{r, results='hide'}
# xor
model_xor_zScored_prior <- brm(
  target ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_prior)
```

```{r, results='hide'}
# xor
model_xor_zScored_comp <- brm(
  target ~ competence + 
    (1 + competence | submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_comp)
```

```{r, results='hide'}
# xor
model_xor_zScored_rel <- brm(
  target ~ relevance + 
    (1 + relevance | submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_rel)
```

```{r, results='hide'}
# some
model_some_zScored_rel <- brm(
  target ~ relevance + 
    (1 + relevance | submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_rel)
```

```{r, results='hide'}
# some
model_some_zScored_prior <- brm(
  target ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_prior)
```

```{r, results='hide'}
# some
model_some_zScored_comp <- brm(
  target ~ competence + 
    (1 + competence | submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_comp)
```

With two predictors:
```{r, results='hide'}
# xor
model_xor_zScored_pri_comp <- brm(
  target ~ prior*competence + 
    (1 + prior*competence || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_pri_comp)
```

```{r}
bayesplot::mcmc_pairs(model_xor_zScored_pri_comp, pars = c("b_prior", "b_competence"))
```

```{r, results='hide'}
# xor
model_xor_zScored_pri_rel <- brm(
  target ~ prior*relevance + 
    (1 + prior*relevance || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_pri_rel)
```

```{r}
bayesplot::mcmc_pairs(model_xor_zScored_pri_rel, pars = c("b_prior", "b_relevance"))
```

```{r, results='hide'}
# xor
model_xor_zScored_rel_comp <- brm(
  target ~ relevance*competence + 
    (1 + relevance*competence || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor_zScored_rel_comp)
```

```{r}
bayesplot::mcmc_pairs(model_xor_zScored_rel_comp, pars = c("b_relevance", "b_competence"))
```

```{r, results='hide'}
# some
model_some_zScored_pri_comp <- brm(
  target ~ prior*competence + 
    (1 + prior*competence || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_pri_comp)
```

```{r}
bayesplot::mcmc_pairs(model_some_zScored_pri_comp, pars = c("b_prior", "b_competence"))
```

```{r, results='hide'}
# some
model_some_zScored_pri_rel <- brm(
  target ~ prior*relevance + 
    (1 + prior*relevance || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_pri_rel)
```

```{r}
bayesplot::mcmc_pairs(model_some_zScored_pri_rel, pars = c("b_prior", "b_relevance"))
```

```{r, results='hide'}
# some
model_some_zScored_rel_comp <- brm(
  target ~ relevance*competence + 
    (1 + relevance*competence || submission_id) +
    (1 | title),
  data = d_critical_zScored_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_zScored_rel_comp)
```

```{r}
bayesplot::mcmc_pairs(model_some_zScored_rel_comp, pars = c("b_relevance", "b_competence"))
```

#### Stats with dichotomized predictors 
Maximal model on raw inference ratings, using anticipated binary categorization of items as predictors 
```{r, results='hide'}
d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = factor(relevance, levels = c(1, 0)),
         competence = factor(competence, levels = c(1, 0)),
         prior = factor(prior, levels = c(1, 0))) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat
# sum code predictors
# low prior: -1, high prior:: 1 
contrasts(d_critical_wide_cat$prior) <- contr.sum(2)
# low comp : -1, high comp : 1
contrasts(d_critical_wide_cat$competence) <- contr.sum(2)
# low rel: -1, high rel: 1
contrasts(d_critical_wide_cat$relevance) <- contr.sum(2)

@@

```{r}
summary(model_xor_cat)
```
We test the following hypotheses for the categorical predictors models (identical for both or and some):
H1: For higher prior of the stronger alternative being true, the inference strength is lower
H2: For higher speaker comptence, the inference strength is higher
H3: For higher listener relevance, the inference strength is higher 
```{r}
# xor
# test H1
hypothesis(model_xor_cat, "prior1 < 0")
# test H2
hypothesis(model_xor_cat, "competence1 > 0")
# test H3
hypothesis(model_xor_cat, "relevance1 > 0")
```

```{r, results='hide'}
model_some_cat <- brm(
  response ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_cat)
```

```{r}
# some
# test H1
hypothesis(model_some_cat, "prior1 < 0")
# test H2
hypothesis(model_some_cat, "competence1 > 0")
# test H3
hypothesis(model_some_cat, "relevance1 > 0")
```

Maximal model on z-scored inference ratings, using anticipated binary categorization of items as predictors 
```{r, results='hide'}
d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = factor(relevance, levels = c(1, 0)),
         competence = factor(competence, levels = c(1, 0)),
         prior = factor(prior, levels = c(1, 0))) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat_zScored
# sum code predictors
# low prior: -1, high prior:: 1 
contrasts(d_critical_wide_cat_zScored$prior) <- contr.sum(2)
# low comp : -1, high comp : 1
contrasts(d_critical_wide_cat_zScored$competence) <- contr.sum(2)
# low rel: -1, high rel: 1
contrasts(d_critical_wide_cat_zScored$relevance) <- contr.sum(2)

@@

```{r}
# xor
# test H1
hypothesis(model_xor_cat_zScored, "prior1 < 0")
# test H2
hypothesis(model_xor_cat_zScored, "competence1 > 0")
# test H3
hypothesis(model_xor_cat_zScored, "relevance1 > 0")
```

```{r, results='hide'}
model_some_cat_zScored <- brm(
  response_centered ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_some_cat_zScored)
```

```{r}
# some
# test H1
hypothesis(model_some_cat_zScored, "prior1 < 0")
# test H2
hypothesis(model_some_cat_zScored, "competence1 > 0")
# test H3
hypothesis(model_some_cat_zScored, "relevance1 > 0")
```

#### Single predictor z-scored models with categorical predictors
For xor:
```{r, results="hide"}
model_xor_cat_zScored_pri <- brm(
  response_centered ~ prior + 
    (1 + prior | submission_id) +
    (1 | title),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_xor_cat_zScored_pri)
```

```{r, results="hide"}
model_xor_cat_zScored_comp <- brm(
  response_centered ~ competence + 
    (1 + competence | submission_id) +
    (1 | title),
  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_xor_cat_zScored_comp)
```
